---
title: "Playing With OpenTargets DATA"
output: github_document
---

```{r setup, include=FALSE}
library(dplyr)
```


```{r }
pathToCSV <- '/home/gummi/netprop/data/interaction/interactionAll.csv'

intData <- read.csv(pathToCSV, stringsAsFactors = FALSE)

dim(intData)

length(unique(c(intData$targetA, intData$targetB)))

```



```{r }

# Filter out any row that have scoring less then 0.75 (perhaps other values should be tested)

intDataFilt  <- intData %>% filter(!(sourceDatabase == "string" & scoring <= 0.4)) # 0.4 only for stringdb

rm(intData)

```


```{r }
# Create an unweighted graph in igraph with targetA and targetB as nodes 
library(igraph)

intGraph <- graph.data.frame(intDataFilt[,c("targetA","targetB" )], directed = FALSE)

# Remove any self loops and redundant edges

intGraph <- simplify(intGraph, remove.multiple = TRUE, remove.loops = TRUE)

# Get number of nodes and edges

length(V(intGraph))
length(E(intGraph))

```

# Read in the assocation data
```{r }

pathToAssocCSV <- "/home/gummi/netprop/data/associationByOverallDirect.csv"

assocData <- read.csv(pathToAssocCSV, stringsAsFactors = FALSE)

types <- table(stringr::str_split(assocData$diseaseId, pattern = "_", simplify = TRUE)[,1])

#    DOID      EFO       GO       HP    MONDO       MP     NCIT      OBI 
#    6876  1420984     5280   117128   846101     2805       28     9836 
# Orphanet     OTAR 
#   119825     2344

```

```{r }

# Filter out only EFO terms

assocDataFilt <- assocData %>% filter(stringr::str_split(diseaseId, pattern = "_", simplify = TRUE)[,1] == "EFO")

length(unique(assocDataFilt$diseaseId))

test_trait<- "EFO_0001073"

# Filter for test trait and remove all rows that have a score less then 0.5

assocDataFiltTest <- assocDataFilt %>% filter(diseaseId == test_trait) %>% filter(score >= 0.5)

rm(assocData)

```

```{r }

# Create the seed vector using the assocDataFiltTest

seedVector <- rep(0, length(V(intGraph)))

seedVector[match(assocDataFiltTest$targetId, V(intGraph)$name)] <- 1

```

```{r }

# Run network propagation with igraph page rank and seed vector

netProp <- page_rank(intGraph, directed = FALSE, damping = 0.85, personalized = seedVector)

# Plot the distribution of the network propagation scores

hist(netProp$vector, breaks = 100)

```

```{r }

# Extract the top 20 genes with the highest network propagation scores

top20 <- sort(netProp$vector, decreasing = TRUE)[1:20]

# Extract the top 20 genes with the highest network propagation scores that are not in the seed vector	

top20NoSeed <- sort(netProp$vector[seedVector == 0], decreasing = TRUE)[1:20]

# Convert the top 20 genes from ensamble to symbol

library(biomaRt)

ensembl <- useEnsembl(biomart = "ensembl", mirror = "useast",dataset = "hsapiens_gene_ensembl")

top20Symbols <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"), filters = "ensembl_gene_id", values = names(top20), mart = ensembl)

top20SymbolsNoSeed <- getBM(attributes = c("ensembl_gene_id", "hgnc_symbol"), filters = "ensembl_gene_id", values = names(top20NoSeed), mart = ensembl)

```

```{r }

# Take the nodes in the top 25% of the network propagation scores and create a subgraph

top25 <- sort(netProp$vector, decreasing = TRUE)[1:round(length(netProp$vector)*0.25)]

subGraph <- induced_subgraph(intGraph, which(netProp$vector %in% top25))

# Perfrom walktrap community detection on the subgraph

subGraphComm <- walktrap.community(subGraph) # only clusters with atelast seed gene

# Do clusters have higher network propagation scores then the rest of the genes in the subgraph

table(subGraphComm$membership)

```

```{r }

# Run clusterprofiler on the genes from the walktrap community detection 

library(clusterProfiler)
library(org.Hs.eg.db)
library(ggplot2)
for( i in unique(membership(subGraphComm))) {
    genes <- names(membership(subGraphComm)[membership(subGraphComm) == i])
    if(length(genes) < 5) {
        next
    }
    ego <- enrichGO(gene         = genes,
                    OrgDb         = org.Hs.eg.db,
                    keyType       = 'ENSEMBL',
                    ont           = "ALL",
                    pAdjustMethod = "BH",
                    pvalueCutoff  = 0.01,
                    qvalueCutoff  = 0.05)

    plot <- barplot(ego) + ggtitle(paste("Cluster ", i)) 
    plot2 <- cnetplot(ego, circular = TRUE, colorEdge = TRUE) + ggtitle(paste("Cluster ", i))	
    ggsave(paste("cluster", i, ".png", sep = ""), plot)
    ggsave(paste("cluster", i, "cnetplot.png", sep = ""), plot2)

}


```





```{r }

# Create a function that generated an AUROC with given a trait and




```



```{r }

# use pROC

library(pROC)

numberOfrepl <- 100

# Compute the AUROC for single trait

proc <- rep(NA,numberOfrepl)

plots <- list()	

for(i in 1:numberOfrepl) {

    # Given a trait, pick 10 random genes that are seeds and remove the trait from the seed vector
    recoverGens <- assocDataFiltTest %>% 
                    sample_n(50,weight = assocDataFiltTest$score) %>% 
                    select(targetId) %>% 
                    unlist() %>% 
                    unname()

    #recoverGens <- sample(V(intGraph)$name,50)

    # Generate a seed vector without the recover genes
    seedVector <- rep(0, length(V(intGraph)))
    seedVector[match(
                        setdiff(assocDataFiltTest$targetId,
                                recoverGens), 
            V(intGraph)$name)] <- 1	

    # Run network propagation with seed vector that doesnt include the recover genes
    netProp <- page_rank(intGraph, directed = FALSE, damping = 0.85, personalized = seedVector)

    # Compute AUROC using the netwerk propagation scores and the recover genes
    recoverGenesVec <- V(intGraph)$name %in% recoverGens

    rocTmp <- roc(response = recoverGenesVec, predictor = netProp$vector,quiet = TRUE)

    proc[i] <- auc(rocTmp)

    #plots[[i]] <- plot.roc(rocTmp)

}

mean(proc)


```



```{r }



avgAUROC(intGraph, seedList, 100, c(0.25,0.75), binarize = TRUE, NormFunc = NULL, settingsForNormFunc = NULL)
avgAUROC(intGraph, seedList, 100, c(0.25,0.75), binarize = TRUE, NormFunc = ECnormalize, settingsForNormFunc = NULL)
settings1 <- list("nSamples" = 100, "perserveDegree" = TRUE, "degreeSampleSmoothing" = 4, "minBucketSize" = 2)
avgAUROC(intGraph, seedList, 3, c(0.25,0.75), binarize = TRUE, NormFunc = permuteTest, settingsForNormFunc = settings1)

```



```


```{r }

# Test the avgAUROC function

seedList <- list(assocDataFiltTest$targetId, assocDataFiltTest$score, assocDataFiltTest$score >= 0.5)

#seedList <- list(sample(V(intGraph)$name,100), rep(1,100), rep(TRUE,100))

#seedList <- list(assocDataFilt$targetId,100), rep(1,100), rep(TRUE,100)

avgAUROC(intGraph, seedList, 10, 0.25, binarize = TRUE)

#x <- c()
#for(i in 1:100) {
#   x[i] <- pROC::auc(pROC::roc(response = sample(c(TRUE,FALSE),100,replace=T,prob=c(0.1,0.9)), predictor = runif(100),direction=">"))[[1]]
#}
#mean(x)


seeds1 <- createSeedVector(intGraph, seedList, binarize = TRUE)
netProp1 <- igraph::page_rank(intGraph, directed = FALSE, damping = 0.85, personalized = seeds1[[1]])$vector
scores1 <- netProp1$vector


ecScore1 <- str(igraph::page_rank(intGraph, directed = FALSE, damping = 1, personalized = seeds1[[1]],algo="arpack"))
scores2 <- scores1/ecScore1$vector

eigeCent <- igraph::eigen_centrality(intGraph, directed = FALSE, scale = FALSE, weights = NULL, options = NULL)

summary(igraph::page_rank(intGraph, directed = FALSE, damping = 1, personalized = seeds1[[1]],algo="arpack")$vector)
```

```{r }

# Run avgAUROC for some EFO traits

read.csv("/home/gummi/netprop/phenotypeMapping.csv", stringsAsFactors = FALSE) -> diseaseMappings

aurocs <- list()

# Filter traits from assocDataFilt that have less then 10 seed genes with score >= 0.5
assocDataFilt10 <- assocDataFilt %>%
                            filter(score >= 0.5) %>%
                            group_by(diseaseId) %>% 
                            filter(n() >= 10) %>%
                            ungroup() 

for(trait in unique(assocDataFilt10$diseaseId)) {

    assocDataFiltTemp <- assocDataFilt10 %>% filter(diseaseId == trait)

    seedsInd <- assocDataFiltTemp$score >= 0.5

    if(sum(seedsInd) < 10) {
        print(paste("Trait ", trait, " has less then 10 seed genes with score >= 0.5", sep = ""))
        next
    }

    seedList <- list(assocDataFiltTemp$targetId, assocDataFiltTemp$score, seedsInd)

    diseaseName <- diseaseMappings[which(diseaseMappings$id == trait),4] 

    aurocs[[diseaseName]] <- avgAUROC(intGraph, seedList, 10, c(0.25,0.5,0.75), binarize = TRUE, weightSampling = FALSE,usePermutedscores = FALSE,permuteSamples = 100)

    print(paste("Finished trait ", diseaseName, " with ", sum(seedsInd), sep = ""))

}

allTraits <- do.call(rbind,aurocs) %>% as.data.frame() %>% write.csv("aurocsNoWeight.csv", row.names = FALSE)

```

```{r }

# Create a function for permuation testing


# Test function

#permuteVals <- permuteTest(intGraph, seedList, binarize = TRUE, nSamples = 100, perserveDegree = FALSE, degreeSampleSmoothing = 3,minBucketSize = 3)
#permuteVals <- permuteTest(intGraph, seedList, binarize = TRUE, nSamples = 100, perserveDegree = TRUE, degreeSampleSmoothing = 1,minBucketSize = 2)
#ssedVec <- createSeedVector(intGraph, seedList, binarize = TRUE)
#settings1 <- list("nSamples" = 100, "perserveDegree" = TRUE, "degreeSampleSmoothing" = 2, "minBucketSize" = 2)
#permuteVals <- permuteTest(intGraph, ssedVec[[1]], netPropScores = NULL, settings = settings1)


```

```{r }
permuteVals <- permuteTest(intGraph, seedList, binarize = TRUE, nSamples = 1000, perserveDegree = TRUE, degreeSampleSmoothing = 2,minBucketSize = 2)

DescTools::Entropy(permuteVals[[1]]) # IS the distribution of the p-values uniform?  LOTS OF 1s, WHY??


```

```{r }

```

```{r }
```



```{r }
library(pROC)
variantCertain = read.csv("/home/gummi/netprop/data/20240101_variantsEFO_MONDO_withSource_updatedJBTS.csv")
variantCertain = unique(variantCertain[,c("diseaseId", "targetId")])

variantCertainMin10 = variantCertain[variantCertain$diseaseId %in% names(table(variantCertain$diseaseId))[table(variantCertain$diseaseId) > 9],]

diseases = unique(variantCertainMin10$diseaseId)


benchmark <-
  function(disease = diseases[1], intGraph = intGraph, phenoDF,prop) {
    variantList <-
      unique(unlist(phenoDF[phenoDF$diseaseId == paste0(disease), 'targetId']))

    if (sum(variantList %in% V(intGraph)$name) != length(variantList)) {
      missingGenes <- length(variantList) - sum(variantList %in% V(intGraph)$name)
      variantList <- variantList[variantList %in% V(intGraph)$name]
      print(paste("Disease ", disease, " has ", missingGenes, " genes that arent in network", sep = ""))
    } 

    testList <-
      sample(variantList, prop*length(variantList))
    
    `%notin%` <- Negate(`%in%`)
    trainList = variantList[variantList %notin% testList]
    
    if (length(variantList) < 10) {
      resAUC = NA
    } else {
      V(intGraph)$targetGene = 0
      V(intGraph)$targetGene[V(intGraph)$name %in% trainList] = 1
      
      pageRank = page_rank(intGraph, personalized = V(intGraph)$targetGene)
      
      pageRankRes = data.frame(
        gene = V(intGraph)$name,
        pageRank = pageRank$vector,
        target = V(intGraph)$targetGene
      )
      pageRankRes$test = 0
      pageRankRes[pageRankRes$gene %in% testList, 'test'] = 1
      
      pageRankRes = pageRankRes[pageRankRes$target != 1, ]
      resROC <- roc(pageRankRes$test, pageRankRes$pageRank,quiet = TRUE)
      resAUC <- resROC$auc
    }
    return(resAUC)
  }

#aucAll <- sapply(diseases, FUN = function(x){benchmark(disease = x, intGraph = intGraphClean)})


# Try running it 10 times for each disease and take the mean
#ellensAUR <- list()
#
#
#for (i in 1:10) {
#    # This was run with prop = 0.9
#  ellensAUR[[i]] <- sapply(diseases, FUN = function(x){benchmark(disease = x, intGraph = intGraph, phenoDF = variantCertainMin10,prop = 0.5)})
#
#}


#unlist(unlist(ellensAUR)) %>% summary()


# Try running it on the data I was working on

resEFO <- sapply(unique(assocDataFilt10$diseaseId), FUN = function(x){benchmark(disease = x, intGraph = intGraph, phenoDF = assocDataFilt10,prop = 0.9)})


```


```{r }


source("/home/gummi/netprop/Code/NetPropFuncs.R")

aurocs <- list()

for(trait in diseases) {

    print(paste("Starting trait ", trait, sep = ""))

    variantCertainTemp <- variantCertainMin10 %>% filter(diseaseId == trait)

    seedsInd <- variantCertainTemp$targetId

    seedList <- list(variantCertainTemp$targetId, rep(1,length(seedsInd)), rep(TRUE,length(seedsInd)))

    aurocs[[trait]] <- avgAUROC(intGraph, seedList, 1, c(0.5), binarize = TRUE, NormFunc = NULL, settingsForNormFunc = NULL)

}

allTraits <- do.call(rbind,aurocs) %>% as.data.frame() 


source("/home/gummi/netprop/Code/NetPropFuncs.R")

aurocs <- list()

for(trait in unique(assocDataFilt10$diseaseId)) {

    print(paste("Starting trait ", trait, sep = ""))

    variantCertainTemp <- assocDataFilt10 %>% filter(diseaseId == trait)

    seedsInd <- variantCertainTemp$targetId

    seedList <- list(variantCertainTemp$targetId, rep(1,length(seedsInd)), rep(TRUE,length(seedsInd)))

    aurocs[[trait]] <- avgAUROC(intGraph, seedList, 1, c(0.9), binarize = TRUE, NormFunc = NULL, settingsForNormFunc = NULL)

}

allTraits <- do.call(rbind,aurocs) %>% as.data.frame() 


summary(allTraits$mean_0.9)
summary(resEFO)

```

# Run ellens code
```



```



```


```{r }

nodeInds <- 1:100





# Create buckets for each of the seed nodes based on their degree
seedDegrees <- degree(intGraph, nodeInds) 

# Get degrees of all nodes in network
allDegrees <- degree(intGraph) 


sapply(b, function(x) sample(x,1))



x <- abs(17- seedDegrees)

``` 


```{r }


# Plot from the auroc.csv file

aurocs <- read.csv("aurocs.csv", stringsAsFactors = FALSE)

library(ggplot2)

# Plot density of values of the AUROC for mean_0.25, mean_0.5, mean_0.75

# Combine the three mean columns into one column and create a new column that indicates which mean column the value came from

aurocsLongDF <- data.frame(vals = c(aurocs$mean_0.25, aurocs$mean_0.5, aurocs$mean_0.75),
                            PercentDropped = c(rep("0.25",nrow(aurocs)), rep("0.5",nrow(aurocs)), rep("0.75",nrow(aurocs))),
                            nSeeds = c(aurocs$nseeds,aurocs$nseeds,aurocs$nseeds))


# plot the density of the mean values

#ggplot(aurocsLongDF, aes(x = vals, fill = PercentDropped)) + geom_density(alpha = 0.5) + theme_bw() + ggtitle("Density of AUROC values for Drop_0.25, Drop_0.5, Drop_0.75")

# Plot violin plots of the mean values

#ggplot(aurocsLongDF, aes(x = PercentDropped, y = vals, fill = PercentDropped)) + geom_violin() + theme_bw() + ggtitle("Violin plot of AUROC values for mean_0.25, mean_0.5, mean_0.75")

# Plot box plots of the mean values side by side 

ggplot(aurocsLongDF, aes(x = PercentDropped, y = vals, fill = PercentDropped)) + geom_boxplot() + theme_bw() + ggtitle("Box plot of AUROC values for Drop_0.25, Drop_0.5, Drop_0.75")

# Plot the relationship between the number of seed genes and the mean AUROC values with a scatter plot and regression line, drop outliers

#ggplot(aurocsLongDF, aes(x = nSeeds, y = vals,color = PercentDropped)) + geom_point() + xlim(0,100) +geom_smooth(method = "lm") + theme_bw() + ggtitle("Scatter plot of AUROC values for Drop_0.25, Drop_0.5, Drop_0.75")


```